## Results

### How do vaccination levels and hesitancy differ by geographic region in the US?
```{r}
plot_vax_by_region()
```
The  faceted histogram above was obtained by combining census data at a county level with the county-level vaccination data from the CDC and state sources. We used the variable "fully_vax" as the data for the histogram and faceted by the census region (South, West, Midwest, Northeast) of the county.

The histogram demonstrates several important ideas. First, the Northeast region of the country displays a clear lead in the percentage of counties that have higher levels of full vaccination among residents. A large majority of counties in the Northeast are at least 65% fully vaccinated, and many are reaching the 70-75% full vaccination levels that can signal the beginnings of herd immunity. Furthermore, the data is more centralized in the Northeast compared to the other regions, suggesting that there is not a large level of hesitancy or belief in misinformation about the vaccine.

Meanwhile, the South is lagging behind the other three regions as a large portion of its counties are still in the 35-45% fully vaccinated range. Additionally, there is wider range in the data, and there are many counties that are still below a 25% full vaccination threshold. This could be a result of or correlated to several influencing factors which will be discussed in other results, including demographics, hesitancy, and politics.

The Midwest and West are very similar in that they both have a peak around the 45-55% range. Both regions display a wide range of vaccination levels in both directions, suggesting that there is a wide array of willingness to get vaccinated within those regions.

```{r fig.align='center'}
library(gridExtra); library(grid); library(lattice)
region_plots <- plot_regions()
region_plots$fully_vax$us
```
We created this map using ggplot's sf geometry, the county-level total population census data, total case data, and total full vaccination data.

The map shows a county-level breakdown of the proportion of the population that is fully vaccinated. Once again, it is clear that there are regions of the country, especially the South (even more specifically, Georgia), that are severely under-vaccinated. In fact, only the Northeast region is reaching higher levels of vaccination, and while there are pockets of high vaccination in the West and Midwest, such as in the Seattle area, those regions do not have a uniformly high vaccination rate thus far. Perhaps most alarming is that despite many counties reaching a 45% to 65% full vaccination rate, only a few select counties have crossed the 65% full vaccination threshold thus far. This is important, as many scientists place the threshold for herd immunity in a community at around 65-70% full vaccination, suggesting that as a country, we are quite far from reaching a safe vaccination level.

```{r fig.align='center'}
region_plots$cases_fully_vax$us
```
This map is colored on a gradient based on the ratio of cases per 100k to the full vaccination percentage. In order to do this, we created a cases/100k variable using a combination of case and census data and then a ratio variable using the cases/100k and the fully vaccinated proportion.

We believed that this variable would be interesting in that it highlights the counties with either high case rates, low vaccination rate, or both. This map suggests that, once again, the South exhibits a much higher (worse) ratio of cases per 100k to vaccination proportion than the rest of the country, while the Northeast is the opposite, with a very low (better) proportion of cases to vaccinations. What is interesting in this map is that most of the West and Midwest also exhibits low ratios. In combination with the previous maps, this suggests that those areas are more average in both cases per 100k and vaccination rate, or have vaccination rates that reflect how bad the pandemic was in the county.

```{r fig.align='center'}
region_plots$hesitant_unsure$us
```

This map displays the county-level percentage of residents who are hesitant to receive the vaccine. There were certainly some surprises in this map. First, Arizona stands out as a state with very high levels of vaccine hesitancy despite being a particularly hard hit state throughout the pandemic and voting for Joe Biden in the 2020 presidential election. Another surprise came in the low hesitancy levels in Nebraska, Utah, North Dakota, and parts of South Dakota, despite those states being very Republican states who voted for Trump in 2020. This is also surprising since much of the South displays the opposite trend despite having similar politics. Thus, there might be other factors at play in these hesitancy levels, potentially race or vulnerability, which will be discussed in later section.

### How is a county's Social Vulnerability Index related to its rate of vaccination?
Here, we plotted county-level vaccination rate distributions by SVI category. The distributions appear approximately normal, with the exception of one intriguing feature: the very high, high, and moderate vulnerability counties all seem to have a bi-modal distribution, with a small peak of very low vaccination levels and a large peak of more moderate vaccination levels. This led us to consider whether there might be a subclass of vulnerable, very low vaccination rate counties in which vaccination rates are influenced by some other variable, a question we explore later.

The second plot, which shows average vaccination rate of counties over time grouped by SVI category, clearly demonstrates that average vaccination levels in higher vulnerability counties have tapered off at lower percentages than those of lower vulnerability counties, with divergence beginning around February 2020. Present-day average vaccination levels for the different SVI categories smoothly increase with decreasing vulnerability. 

Given that the histograms for distribution of vaccination rates of broadly differing SVI categories overlap considerably, we wanted to determine whether the present difference in average vaccination rates between vulnerable (Very High, High, and Moderate Vulnerability SVI categories) and non-vulnerable (Low and Very Low Vulnerability) observed in the plot of vaccination rates over time was statistically significant, so we performed a one-sided t test with the alternative hypothesis that vulnerable counties have a lower average vaccination rate than non-vulnerable counties. The results were very significant, with a p-value near 0. 

```{r}
plot_svi()
plot_vax_over_time()
```


### How are the prevailing political views of a county related to its rate of vaccination?
We plotted total percentage of county votes cast for Donald Trump in the 2020 presidential election versus total percentage of adult population vaccinated. As anticipated, there appears to be a definite negative correlation between the percent of the county that voted for Donald Trump in 2020 and the percent of the population that is vaccinated. Donald Trump and other Republicans have consistently shown themselves to be less in favor of lockdown orders, mask mandates, and COVID-19 vaccinations than Democrats, so the counties in which a larger percentage of the votes were cast for Donald Trump in the 2020 presidential election are likely to be those for whom these anti-COVID precaution sentiments seemed reasonable. However, there are outlier counties with far more or fewer people vaccinated than would have been predicted based solely on their political views. In particular, there was a cluster of Democratic counties with unexpectedly low vaccination rates. The next section explores race as a potential contributing factor predicting COVID-vaccination rates in outlier counties.

```{r}
plot_trump_vacc()
```


### How does the percentage of Black residents in a county relate to its level of vaccination?
#### What are the racial demographics of counties that do not conform to the expected relationship between politics and vaccination?

There appeared to be a distinct cluster of democratic counties with unexpectedly low vaccination rates in the lower left corner of the plot in the previous section. We hypothesized that these may be counties with high percentages of non-white residents. To investigate these and other outlier counties, we performed a robust linear regression (Huber regression) on the relationship between proportion of votes cast for Trump and proportion of population vaccinated. We then examined both Democratic (<40% of all votes in the 2020 presidential race cast for Trump) and Republican (more than 60% of all votes cast for Trump) in the top 10% (unexpectedly high vaccination rates) and bottom 10% of residuals (unexpectedly low vaccination rates) from this linear regression. We produced histograms of the proportion of Black residents in each of these four categories of counties. For the counties with unexpectedly high vaccination rates (both Democratic and Republican), the distribution of proportion Black residents peaked near 0. For the low-vaccination Republican counties, many counties also had a proportion of Black residents near 0, but there was another peak around 0.25. For the Democratic low-vaccination rate counties, the results were even more striking: a majority of counties had more than 35% Black residents. Clearly, counties with lower vaccination rates than predicted by their political leanings are much more likely to have more Black residents than counties with higher vaccination rates than predicted by their political leanings. 

Next, we performed an ANOVA examining the mean percent Black for all four groups of outlier counties (low vaccination, Democrat; low vaccination, Republican; high vaccination, Democrat; high vaccination, Republican) followed by a Tukey Test pairwise comparison of means. All pairs showed significant difference in means at significance level 0.01 except high vaccination, Republican versus high vaccination, Democrat, demonstrating that the racial disparities visible in the histograms are statistically significant. 

```{r}
rnames <- tibble::tribble(
	~ "rn", ~ "Comparison",
	"bottom_red-bottom_blue", "Bottom 10% GOP vs bottom 10% Dem",
	"top_blue-bottom_blue", "Top 10% Dem vs bottom 10% dem", 
	"top_red-bottom_blue", "Top 10% GOP vs bottom 10% Dem", 
	"top_blue-bottom_red", "Top 10% Dem vs bottom 10% GOP", 
	"top_red-bottom_red", "Top 10% GOP vs bottom 10% GOP", 
	"top_red-top_blue", "Top 10% GOP vs top 10% Dem"
)

tk <- race_hist_plot()$orig_df %>% 
	tibble::as_tibble(rownames = "rn") %>%
	dplyr::inner_join(rnames, by = "rn") %>% 
	dplyr::select(Comparison,
								"Difference in Mean Proportion Black" = diff, 
								"95% Confidence Lower Bound" = lwr, 
								"95% Confidence Upper Bound" = upr, 
								"p-value" = `p adj`)

kable(tk,
			digits = 4) %>% 
	kableExtra::kable_styling(full_width = T)
```

#### What are the racial demographics of counties with high levels of vaccine hesitancy and low vaccination rates?
Because political values have already been identified as a predictor of vaccination, we decided to filter by percent voting for Donald Trump before examining race versus vaccine hesitancy. Additionally, we were particularly interested understanding the factors controlling vaccination in the previously identified cluster of low-vaccination Democratic counties, so we chose to examine only counties in which less than 40% of the votes in the 2020 presidential election were cast for Donald Trump. Finally, because there are a large number of counties with close to 0% Black residents and these are not relevant to our examination of vaccine hesitancy in Black communities, we limited the counties to only those with a Black population of 5% or greater. Once we had selected these counties, we plotted the estimated percentage of the population that was hesitant or unsure about the Covid vaccine versus the percentage of the population consisting of Black residents and conducted a linear regression. Ultimately, we sought to investigate whether there was a correlation between the percent of the population that was Black and levels of vaccine hesitancy in Democratic counties. 

We found that among Democratic counties with more than 5% Black residents, vaccine hesitancy was indeed correlated with race. This suggests our previous observation that undervaccinated counties are more likely to have larger Black populations may result from greater reluctance of residents in these counties to accept the vaccine, rather than stemming from limited vaccine supply and distribution disparities, although the latter possibility cannot be ruled out.

```{r}
plot_vacches_race()
```

#### What are the racial demographics of high SVI counties with low vaccination rates?
Because of our previous observations about the relationship between a county's proportion Black residents and its levels of vaccination and vaccine hesitancy, we hypothesized that the group of high SVI counties with especially low vaccination rates that we observed in the histograms of vaccination levels for counties in differing SVI categories might consist of counties with large Black populations. We assessed this theory by constructing two sets of violin plots showing distributions of percent Black residents based on SVI category and vaccination levels. The first set of violin plots shows counties in the "Very High" "High" or "Moderate" SVI categories and compares low vaccination counties (less than 15% vaccinated) versus all other counties, while the second set shows the same comparison but for "Low" or "Very Low" SVI categories. While the low vulnerability counties show similar distributions of percent Black residents for both the low and the high vaccination rate groups, for high vulnerability counties, the distribution is strikingly skewed towards more Black residents for the low vaccination rate counties in comparison with the high vaccination rate counties.
```{r}
make_svi_race_violinplot(1, 0.4)
make_svi_race_violinplot(0.39, 0)
```

### Did the severity of the pandemic within a county affect residents' desire to get vaccinated?

```{r fig.align='center'}
region_plots$deaths_fully_vax$us
```

This map is colored based on the ratio of deaths per 100k to vaccination rate rather than the cases per 100k to vaccination rate. While this is a rudimentary form of analysis, there do seem to some differences in the two ratios, especially in the mountainous states of the West, such as Idaho, Wyoming, and Utah. Specifically, those states had a much lower ratio of deaths to vaccinations relative to their cases to vaccinations ratio, which might suggest that in those areas there might be some correlation, although there is nothing definitive. To further analyze this possibility, we plotted deaths per county vs. vaccination rate:

```{r}
plot_deaths_pc_vaccine()
```

As far as we can tell from this chart, there is no overall relation between severity and vaccination rates. However, we haven't yet completed any regressions, so all we have is the plot. We want to regress controlling for demographics and vaccine hesitancy.